---
title: "Solutions to the potential for unintended conservation consequences of MPAs"
subtitle: Conditions for success and unintended consequences in pelagic MPA design
author:
    - Daniel Ovando
    - Darcy Bradley
    - Lennon Thomas
    - Echelle Burns
date: "`r Sys.Date()`"
output: 
  bookdown::pdf_document2: default
  bookdown::html_document2: default
bibliography: references.bib
params:
  run_name: ["v0.5"]
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, include = TRUE)

source(here::here("scripts", "00_setup.R"))

theme_set(marlin::theme_marlin())

```

```{r load-results}

results_files <- list.files(file.path(results_path))

rdses <- results_files[str_detect(results_files,'.rds')]

walk(rdses,~assign(str_remove(.x,".rds"), read_rds(file.path(results_path,.x)), envir = .GlobalEnv)) #load and assign all .rds objects in the appropriate results folder

rdatas <- results_files[str_detect(results_files,'.Rdata')]

walk(rdatas, ~load(file.path(results_path,.x), envir = .GlobalEnv))

```

\*\* something to point out: increased catches are a paradox, either
come from rebuilding stock or concentrated fishing, but trajectory can
tell you something. not possible for MPAs to increase catches without
decreasing catches first if they are rebuilding? \*\*

Seasonal movemetn is kind of interesting, basically just dilutes the
extremes

less resilient species are both more in need of protection and easier to
accidentally fish harder

# Introduction

<!-- Empirically, how do you differentiate increase in catch from fishing harder vs increases in catch from rebuilding: rebuilding has to have a decrease at first -->

Marine protected areas (MPAs) are a growing part of the marine
conservation toolbox, with 5.9% of the world's oceans currently in
implemented MPAs [@grorud-colvert2021], and bodies such as the IUCN
calling for 30% protection by 2030. The modern science of MPAs was
largely developed around coastal ecosystems, often characterized by
relatively static features such as coral reefs [citation]. However,
large MPAs are increasingly being implemented and considered across
pelagic ecosystems. These pelagic systems are characterized by highly
mobile species and fishing fleets using a diverse mix of static and
dynamic ocean features [@game2009; @lubchenco2003]. These ecosystems are
also home to combinations of relatively abundant species targeted by
fishing, such as many tuna stocks [@pons2016], and highly threatened
species caught as incidental "bycatch" in targeted fishing operations,
such as oceanic whitetip sharks [@pacoureau2021]. Large pelagic MPAs are
increasingly being proposed specifically to benefit these by-catch
populations while minimizing losses or where possible providing benefits
to fisheries [citation].

In this paper we evaluate the potential conservation and fishery effects
of MPAs within the pelagic context. We focus in particular on the
potential for pelagic MPAs to provide conservation benefits to bycatch
species. While the effect of MPAs on fisheries yields is a continuously
debated topic, the conservation impacts of MPAs are seemingly simpler,
albeit challenging to detect in many cases [@ovando2021a], due in part
perhaps to the perceived obvious nature of their benefits: if you
prevent fishing of a species within an area, there will be more of that
species in that area, and so the conservation benefits of MPAs will
increase with MPA size. \@hastings2017 clearly illustrates this idea.
Using a simple two-species two-patch analytical model, they derive
conditions under which MPAs can solve the "weak-stock" problem in which
a less productive species is caught as btcatch by a fleet targeting a
more valuable and productive stock.

However, as @hastings2017 point out, their results depend on a very
specific set of assumptions, namely that 1) adults are perfectly
sedentary 2) the target and bycatch species overlap perfectly in terms
of habitat and 3) recruits are perfectly mixed throughout the system.
Pelagic species are likely to violate one or more fo these assumptions.
We build on \@hastings2017 to consider the role of MPAs in reducing
bycatch while managing tradeoffs with fisheries under a more resolved
model reflecting the dynamics of pelagic ecosystems and fisheries.

Pelagic ecosystems present both opportunities and challenges for MPAs.
The highly productive and mobile nature of targeted fisheries in these
systems may mean that highly targeted MPAs can be designed that provide
effective conservation at minimal fishery costs. Strategic protection of
key aggregation locations may enhance depleted pelagic fisheries [xx
that Indian ocean study I can't remember right now]. However, the
dynamic nature of pelagic systems also creates a risk of unintended
conservation consequences of MPAs. When an MPA is implemented on former
fishing grounds, fishing fleets have two choices: exit the fishery or
move to new fishing grounds. When MPAs redistribute effort away from one
area and into another, they can accidentally increase fishing effort on
species found within the new fishing grounds [@abbott2012]. This problem
of policy "leakage" is well known in many environmental fields
[@citations] but the general properties of this problem have received
relatively little attention in the theoretical or empirical MPA
literature [see
@abbott2012;@savina2008;[fulton2011a;\@baum2003](mailto:fulton2011a;@baum2003){.email}
for notable exceptions].

This study addresses this gap using a novel bio-economic simulation
model, capable of reflecting several of the key challenges of pelagic
ecosystems, namely high and variable movement rates, mobile fishing
fleets, and heterogeneous species distributions. We first provide a set
of stylized simulations intended to illustrate the potential and
challenges of achieving conservation objectives through pelagic MPAs,
and highlight particularly important drivers and the magnitude and
direction of MPA effects. We then present a more focused case study
based on the Western and Central Pacific Ocean, examining the potential
outcomes of pelagic MPAs on bycatch and target species.

<!-- We fill this gap by examining the general range of conditions… -->

\<!-- \# Assessing the conditions under which this may be a problem
using a generalized simulation environment

General insights about design strategy, effect of varying degrees of
habitat overlap, etc, tradeoffs with fisheries

-   Given that "it depends" nature of the problem, assess the specific
    potential for this problem in a case study

    -   specific insights about design strategy, effect of varying
        degrees of habitat overlap, etc, tradeoffs with fisheries

Our goal here is not to provide tactical recommendations on the design
of blue-water MPAs, but rather to provide a strategic evaluation of the
sensitivity of bycatch populations and targeted fisheries to key
unknowns and design choices in MPA planning. --\>

<!-- The outcome of MPAs for fishing and conservation are highly sensitive to -->

<!-- a wide range of economic and ecological factors. In order to constrain -->

<!-- the possible parameter space, we tune a simulation model based around -->

<!-- best available knowledge of the species and fleets operating in the -->

<!-- areas of the western central pacific ocean (WCPO) while allowing key -->

<!-- unknowns to vary, in particular the nature of adult and larval movement, -->

<!-- the degree of habitat overlap between different species, the economic -->

<!-- incentives facing fishing fleets, and the design strategy used in MPA -->

<!-- placement XX. Most conventional MPA design models are either static, -->

<!-- single species, or both. We present a dynamic, multi-species simulation -->

<!-- model that allows us to cohesively assess the impact of MPAs on a suite -->

<!-- of by-catch and targeted species collectively affected by multiple -->

<!-- fishing fleets.  -->

<!-- We show - something about how when species are homogeneous in space, -->

<!-- conservation gains are basically guaranteed except for constant catch -->

<!-- scenearios, which need to be considered in tuna-type scenarios - When we -->

<!-- introduce heterogeneity in stock structure roughly reflective of -->

<!-- distributions in the WCPO we do / do not begin to see tradeoffs in -->

<!-- conservation outcomes among bycatch species - quantify if there is a -->

<!-- concave function to this - Specific design paradigms are / are not more -->

<!-- / less resilient to this issue - Compare tradeoffs between MPA designs -->

<!-- that minimze internal bycatch tradeoffs (there have to be none if MPA = -->

<!-- 100%) and fishery outcomes -->

<!-- We show that MPAs increased biomass of by-catch species in XX% of -->

<!-- simulations, and decreased them in XX% of simulations. On average a 1% -->

<!-- increase in by-catch species resulted in a XX% change in fishery -->

<!-- catches... Outcomes were most sensitive to XX and XX, and least -->

<!-- sensitive to XX and XX. We do not suggest that these same results would -->

<!-- hold for other systems, but rather demonstrate the importance / -->

<!-- resilience of MPAs to key uncertain variables. In particular for pelagic -->

<!-- ecosystems, our work demonstrates the importance of considering the -->

<!-- collective impacts of space, vulnerability, life history, and fleet -->

<!-- dynamics in MPA design for conservation. -->

# Methods

@hastings2017 employed an analytically tractable model to derive
conditions under which MPAs would benefit bycatch species. However they
also pointed out that even small increases in model complexity quickly
make analytical solutions to the effects of MPAs intractable. As such,
we employ a numerical simulation strategy in the form of a spatially
explicit, multi-species, multi-fleet, bio-economic model we call
[`marlin`](https://github.com/DanOvando/marlin). Models of MPAs span a
wide range of complexity, from simple analytical tools to end-to-end
ecosystem models [@fulton2015]. `marlin` follows a "models of
intermediate complexity" (MICE) philosophy [@plagányi2014] targeted
specifically around the role of technical interactions across species.
Many MPA models simulate the impacts of MPAs on a single-species level
[@ovando2021a]. This can be useful for for example examining the ways in
which species-specific life history traits interact with fleet dynamics
to influence MPA outcomes. However, bycatch problems are driven by
technical interactions, in which fishing effort for a target species
incidentally captures another species. `marlin` accounts for this by
allowing the user to simulate an arbitrary number of fleets capable of
catching an arbitrary number of species in a spatial area. Different
fleets can place different economic value on different species, while
also having species-specific selectivity functions. This allows us to
simulate bycatch dynamics of MPAs in pelagic ecosystems.

## `marlin`

### Fleet Dynamics

The key feature of `marlin` for the purposes of this project is the
ability to capture technical interactions across species. While marlin
does not simulate species interactions, fishing fleets in marlin are
capable of targeting and affecting multiple species simultaneously. This
allows us to capture for example the potential impacts of displacement
of effort from one area into another across multiple species.

A fishing fleet or *metier* is defined in `marlin` by a total pool of
effort that can have different catchability, prices, and selectivity for
individual species within a simulation. For example, both a longline and
purse seine fleet may capture bigeye (*Thunnus obesus*) and skipjack
(*Katswamus pelamus*), but the longline fleet is much more likely to
catch bigeye than skipjack, and *vice versa*. The fishing fleet then
generates catches, revenues, and profits from individual species that it
targets, and makes decisions based on its total objective function
across all species.

Revenues *R* for fleet *f* in time step *t* are given by

$$
 R_{t,p,f} = \sum_s^S \sum_a^A \Pi_{s,f}\frac{\alpha_{a,f}q_{s,f}E_{t,p,f}}{F_{t,p,s,a}} \times Y_{t,p,s,a}
$$

where *p* is patch, *s* is species, *a* is age, $\alpha$ is selectivity
at age, *q* is catchability , *F* is total fishing mortality, and *Y*
are total catches.

$$
Y_{t,p,s,a} = \frac{F_{t,p,s,a}}{z_{t,p,s,a}} \times b_{t,p,s,a} \times (1 - e^{-z_{t,p,s,a}}) 
$$

where *F* is total fishing mortality, given by

$$
F_{t,p,s,a} = \sum_f^F\alpha_{a,f}q_{s,f}E_{t,p,f}
$$

and

$$
 z_{t,p,s,a} = F_{t,p,s,a} + m_{s,a}
$$

Where *m* is natural mortality.

This means that each species experiences a total mortality at age in a
time step in a patch, individual fractions of which are portioned off as
catches and subsequently revenues for each fleet. Given revenues, we
then calculate profits \$\\phi\$ per

$$
\phi_{t,p,f} = R_{t,p,f} - c_f(E_{t,p,f} + 1)_{t,p,f}
^{1.3}$$

where *c* is cost per unit effort and, the 1.3 exponent is in the manner
of @costello2016a . The addition of one to the total effort is to
constrain the value to be greater than 1, preventing undesired behavior
where fishing effort actually gets cheaper if total effort is less than
1.

The important thing to note here then is that the revenues and profits
for a given fleet are a function of the prices and catchability of
individual species, but that mortality itself is independent of price.
So, fleet that has selectivity and catchability greater than zero can
exert fishing mortality on a species even if it obtains no value from
that species.

The fleet then decides how to allocate its effort in space based on one
of four possible spatial allocation strategies. The ideal free
distribution (IFD) (citation) is the standard method for distributing
fishing fleets in spaces. However, the IFD presents a number of
complications for our study. First, in order to simplify the assessment
of the impacts of different fishing mortality rates we simulate
scenarios with a fixed total pool of effort. As MPAs change the spatial
distribution of profits though, the total amount of effort required to
satisfy the IFD may not equal the total amount of effort required by the
model. Second, even if we were to implement the IFD for a single fleet,
the model is capable of tracking multiple fleets at once. This means
that the IFD for a single fleet would be a game-theoretic outcome of the
actions of each of the individual fleet conditional on the actions of
all other fleets. While possible to implement this would slow down our
model runs to the point of practically preventing large-scale evaluation
of the effects of MPAs on bycatch.

As such we explored a series of "next best" fleet distribution
algorithms. While not the IFD in any individual time step, over time
they start to approximate the IFD (check this) as each assumes that
fleets base their decisions for the current time step on the outcomes in
the prior timestep, meaning that the impacts of the actions of other
fishing fleets are somewhat internalized. The four possible fleet
distribution algorithms are

1.  Revenue per unit effort (RPUE): The fleet distributes itself in
    space based on revenue per unit effort
2.  Revenue: The fleet distributes itself in space based on revenue
3.  Profit per unit effort (PPUE): The fleet distributes itself in space
    based on profit per unit effort
4.  Profit: The fleet distributes itself in space based on profit

### Population Model

Lots of equations. Pretty standard age-structured population models,
extra bells and whistles that need explaining

-   Design of movement and habitat

    -   Movement is a product of distance and habitat, such that
        critters will all else equal move to the closest patches with
        the best habitat

-   Ontogenetic shifts

-   Nature of fleet tuning (shared effort different q per species)

Fleets exert a given amount of effort in a patch each time step. The
catch resulting from that effort is given by the Baranov [@baranov1918]
equation

$$C_{f,s,p,t} = \frac{q_{f,s}E_{f,p,t}}{q_{f,s}E_{f,p,t} + m_s} \times \sum_{a=0}^{A}B_{s,p,t,a} \times (1 - e^{-(q_{f,s}E_{f,p,t}S_a + m_s)}) $$

Where $C$ is catch by fleet $f$ of species $s$ in patch $p$ in timestep
*t*. $E$ is total effort by that fleet in a patch and time step, $m$ is
natrual mortality for a given species, $B$ is biomass at age *a*, and
$S$ is selectivity.

$q$ is the catchability coefficient for fleet $f$ and species $s$.

Fleets distribute themselves in space according to revenue per unit
effort

$$R_{f,p,t} = \frac{\sum_{s=1}^{S}v_{f,s}C_{f,s,p,t}}{E_{f,p,t}}$$ Where
$v$ is the price per kg of each species. Bycatch in the model is defined
through the price. An unregulated species with no commercial value would
have $v = 0$, whereas a regulated and prohibited species might actually
have a negative price.

Effort per fleet is set a a constant of 1 per patch, and the model then
tunes $q_{f,s}$ such that given all the other parameters of the model
the fishing mortality rate for each species matches the best available
level for that species. This is what allows the model to capture bycatch
dynamics: A fleet can distribute itself purely based on revenues from
target species, but through the $q$ parameter still exert fishing
mortality on other species, even if that species has a price of zero, if
those bycatch species inhabit the same patches as the more valuable
target species.

### 

### MPA Design Strategies

MPA design is governed by four general parameters: the number of cells
to place inside an MPAs, and the design strategy used to place the MPAs,
the species considered in the MPA planning process, and the accuracy of
the information used in the planning process.

We consider `r n_distinct(experiment_obj$placement_strategy)` different
MPA placement strategies.

For the "depletion" strategy, MPA locations are prioritized based in
proportion to the spawning stock biomass (SSB) of the most depleted
species. So, patches with high amounts of highly depleted species are
prioritized over patches with high biomass of lightly depleted species.

For the "rate" strategy, MPAs are placed in based on the
depletion-weighted catch relative to the total catch in a patch. So,
patches where with high rates of catch of depleted species relative to
totcal catch are prioritized.

For the "avoid_fishing" strategy, MPAs locations are prioritized
inversely to fishery catches. So, patches with low catches are
prioritized over patches with high catches.

For the "target_fishing" strategy, MPAs locations are prioritized
proportionally to fishery catches. So, patches with high total catches
are prioritized over patches with high catches.

For the "area" strategy, we simply assign patches in sequential order
until the desired total MPA size is reached.

The information used to inform each MPA design strategy (except area)
can include a multiplicative observation error term $\sigma_{obs}$ where

$$information \sim lnorm(data,\sigma_{obs}) $$

For example, for the depletion scenario, $\sigma_{obs}$ controls the
accuracy of the spatial depletion data used to prioritize MPA placement.

## Generalized Experiments

Given the complexity of the operating model used here, a complete
sensitivity analysis across all parameters is not practical. Instead, we
focus on the role of life history, fishing pressure, and MPA design.

### Species & Fleets

Each simulation considers `n_distinct(marlin_inputs$scientific_name)`
species comprised of the top 12 species by landings reported in the tuna
fisheries of the Western and Central Pacific Ocean (WCPO). Rough Life
history parameters are pulled from `FishLife` [@thorson2020]. For
teleost fishes, we assumed that fecundity was proportional to
weight-at-age times the maturity ogive. For cartilaginous species, we
adopt a fixed fecundity relationship of XX pups per mature individual XX
check on what I actually implemented here XX. The unfished spawning
stock biomass for each of the included species was tuned in the model to
reflect best available estimates from stock assessments.

Each simulation includes a longline and a purse-seine fleet. Each
species can have a separate price (USD/kg) and selectivity ogive for
each fleet, as dictated by the literature.

### Species Overlap

We simulate degrees of species overlap using two parameters, $\sigma_c$
and $\sigma_h$. All species start with the assumption that the centroid
of their preferred habitat occurs in the center-most patch of teh two
dimensional system simulated by marlin. For our simulations, we then
generated a random deviation from that default centroid for each species
using $\sigma_c$ in the manner

$$centroid_s \sim normal(center,\sigma_c)$$ where *s* denotes species,
and *center* is the coordinate of the central cell. In this way, the
larger $\sigma_c$ grows, the more variation there is betwen the centroid
of the preferred habitat for each species.

We then control the spread of a creatures preferred habitat around that
centroid per a Gaussian dispersal kernal in the manner

$$habitat_{p,s} = \frac{1}{\sqrt{2\pi\sigma_h^2}}e^{-\frac{d_p^2}{2\sigma_h^2}}$$

Where $d$ is the euclidian distance from the $centroid_s$ to patch $p$
and $\sigma_h$ controls the dispersal of the habitat around that range.

### Seasonal Migration

For this paper `marlin` is set to run with two seasons per year. When
seasonal movement is turned on, animals use one habitat matrix in season
one, and the inverse of that habitat matrix in season 2.

## Case Study Application

In order to constrain the parameter space somewhat, we also ran a series
of case study experiments. Rather than iterating over different habitat
configurations and movement dynamics, we defined habitat based on
reported catch per unit effort (CPUE) from the WCPO. XX more methods
here XX

### MPA Network Optimization

For the case study section, as a point of comparison we calculate an
optimized MPA network along with the "rule of thumb" networks based on
the design strategies outlined in section XX. The optimization routine
is based on an objective function of the form

$$obj = \alpha \times scale(\sum_{s=1}^{S}\frac{ssb_s}{ssb0_s}) + (1 - \alpha) \times scale(\sum_{s=1}^{S}R_s) $$

where $\alpha$ controls the relative weight assigned to a "conservation"
objective based on depletion, relative to an "economic" objective based
on revenues. The 'scale' function rescales both objective components to
the range of 0 to 1 to facilitate interpretation of $\alpha$ xx come up
with a better version of this XX

The optimization routine itself works through a
sampling-importance-resampling algorithm. In the first iteration of the
algorithm, the marginal contribution of each individual patch to the
objective function is calculated. The patch with the highest objective
function value is then placed in an MPA. For the remaining iterations,
we sample *X* cells from the remaining non-MPA cells with probability
proportional to their marginal value to the objective function in the
last iteration. We then calculate the marginal value of each of these
cells when individually added to the MPA network begun in the prior
step. The cell with the higheset marginal value is then added to the MPA
network, and the marginal values of the non-selected but sampled cells
are updated based on the values in that iteration. The process then
continues until the desired MPA size is reached.

## Assessing Performance

Each model run started at equilibrium conditions and was then run for
`r years` years (`r seasons * years` seasons. We assessed performance by
comparing relevant metrics (e.g. depletion and revenue) under a given
scenario with MPAs relative to the values of those same metrics in the
simulated counterfactual world identical in every way save for the
presence of MPAs.

# Results

what we found was a series of neccesary but not sufficient conditions 

make the tautology clear: illustrating a point, not proving a thing since clearly depends on if you assume displacement, and then degree of non-overlap etc. 

## Generalized Experiments

We focused first on the range of impacts that MPAs can have on mature
biomass across the range of variables explored. To illustrate, we pulled
out two example simulations, one where MPAs provide relatively large
benefits across all three species, and a second where MPAs actually
produced substantial losses in some species. In the "best case"
scenario, each of the three species shares a common habitat
distribution, and so protection of the central habitat of one species
results in protection for all species. In the "worst case" scenario,
each of the species have markedly different centroids. An MPA designed
primarily around the tuna and marlin populations then ends up displacing
effort into the shark habitat, producing benefits for the first two
species but driving the shark population down.

```{r fig-info, fig.cap="Illustration of ways in which variables can interact to affect MPA outcomes."}

knitr::include_graphics(file.path(results_path,"fig_info.pdf"))
```

the point is not once answer

Looking at the full
`r n_distinct(total_experiment_results$placement_id) * n_distinct(total_experiment_results$state_id)`
experiments run, the results were highly variable, with MPAs capable of
producing substantial conservation gains to substantial conservation
losses and everything in between. 


Design matters size matters species matters habitat matters
```{r}

compact_experiment_results %>%
  filter(prop_mpa < 1) %>%
  # filter(f_v_m > 1.5) %>%
  group_by(state_id, placement_id) %>%
  ggplot(aes(
    prop_mpa,
    pmin(2,biodiv / bau_biodiv -1),
    color =  sigma_habitat,
    group = interaction(placement_id, state_id)
  )) +
  geom_line(alpha = 0.25) +
  geom_hline(aes(yintercept = 0), linetype = 2, color = "red") +
  facet_grid(critter ~ placement_strategy) +
  scale_color_viridis_c(name = "Habitat Difference", guide = guide_colorbar(frame.colour = "black",ticks.colour = "black")) +
  scale_y_continuous(labels = scales::percent,"Change in SSB/SSB0 relative to BAU") +
  scale_x_continuous(labels = scales::label_percent(accuracy = 1), name = "MPA Size") +
  theme(legend.position = "top")

```




```{r}

critter_results <- compact_experiment_results %>%
  group_by(state_id, placement_id) %>%
  mutate(real_sigma_centroid = sd(centroid)) %>%
  mutate(delta_biodiv = pmin(2,biodiv / bau_biodiv - 1),
         loss = biodiv < bau_biodiv,
         rec_form = as.factor(rec_form)) %>% 
  mutate(critter = fct_relevel(critter,c("Tuna", "Marlin","Shark")),
         rec_form = fct_recode(rec_form, "Global DD, Rec. By Habitat" = "0", "Local DD, local Recruitment" = "1", "Local DD, Disperse Recruits" = "2", "Disperse Larvae, Poset-Settlement DD" = "3"))

reg_2 <-   rstanarm::stan_glm(
  loss ~   f_v_m  + critter + seasonal_movement + adult_movement_sigma + rec_form + recruit_movement_sigma + prop_mpa + sigma_habitat + spatial_allocation,
  family = binomial(),
  data = critter_results %>% filter(between(prop_mpa,.2,.4), placement_strategy == "area"),
  cores = 4,
  chains = 4
)

fig_11 <- plot(reg_2)

fig_11

```



We do not focus into the individual
effects of particular variables as we found them to be highly dependent
on the exact question asked and the set of data included, and since we
cannot ascribe any actual likelihood to any individual state of nature
simulated here.

However, three important features do emerge. The first is that life
history plays an important role in the potential for unintended
conservation consequences from MPAs. All else equal, MPAs were much less
likely to produce losses in spawning biomass for our tuna-like species
than the marlin or shark-like species. Given that within the broad life
history traits we do vary adult and larval movement, fishing mortality,
as well as the timing of density dependence, we believe this result is
mostly due to the high resilience (high steepness, low age at maturity,
fast growth) of the tuna-like species relative to the marlin and
shark-like species.

The second is that design matters...

```{r}

```


The third is that extreme outcomes (both positive and negative),
particularly at lower MPA sizes, are largely driven by the presence of
small range sizes interacting with seasonal migrations. Across species,
simulations without seasonal migrations and small ranges were much more
likely to produce extreme results than those with seasonal migrations
and/or larger ranges. Presence of seasonal migrations acts as a
dilutant: As a given static MPA only provides protection, or displaces
fishing effort, onto a particular area in only half of the year, both
the benefits and costs of MPAs are smaller. When seasonal migrations are
not present and the species range is small, MPAs are more capable of
providing either large benefits or large losses.

\\Big message: even constrained by the same initial depletion and core
biology (e.g. movement rates), a wide range of outcomes are possible for
each species for a given MPA size, dependent on the design strategy,
habitat overlap, migrations, quality of information, etc. Across the
board as MPA size increases the ratio of positive to negative outcomes
increases, until all species achieve their maximum conservation
potential in every iteration at 100% MPA. HOwever, every species
experienced a set of simulations, roughyl xx%, that resulted in net
losses in SSB/SSB0 relative to BAU. Losses were smaller than potential
gaines, but reached meaningful levels in some species, and interestingly
peaked in many cases right before max MPA (reflecting concentration of
fishing in small amounts of habitat in habited by the species).Important
to stress in here that there are no probabilities in this, no way of
knowing which state is more probable than others, just given the range
of things simulated more ways for a bigger positive effect. Also make
clear though, biggish losses (10% or more) possible, but rare.

## Case Study

The general experiments conducted show two general insights

1.  Even conditional on fishing exploitation and general life history
    strategy, the effects of MPAs on a group of species with technical
    interactions is highly variable and dependent on habitat
    distribution, design decisions, and data quality. xx make this more
    explicit.

2.  Some degree of unintended conservation consequences occured across
    nearly all of our simulations below MPA coverage near 100%, though
    small losses were much more common than large. Different design
    strategies are better suited to avoiding unintended consequences.

There two general results beg some important questions. First, are
unintended consequences inevitable given the kinds of technical
interactions explored here, or can optimized MPA networks reduce or
eliminate these unintended consequences. Second, one clear solution to
the problem of unintended consequences is to simply make the MPAs large:
if MPAs cover 100% of the area there can be no direct unintended
conservation consequences. The potential for negative impacts on
economic fishery outcomes increases as MPAs reach large size though.
What then are the tradeoffs between reducing the potential for
unintended conservation consequences and fishery revenues?

To address these questions, we ran a more limited "case study" set of
simulations, based on the dynamics of the WCPO tuna fisheries. Rather
than iterating over different movement dynamics and habitat
configurations, we fix habitat layers based on reported CPUE for the
species in question, and xx other things we fix xx.

For this subset of simulations then, along with the various "rule of
thumb" MPA design strategies we ran an optimization routine (see section
XX) across a range of objective function weightings, in the manner of
@rassweiler2014. This allowed us to assess 1) the degree to which
unintended conservation consequences of MPAs can be reduced by optimized
design and 2) the magnitude of tradeoffs between conservation outcomes
and fishery revenues under both rule-of-thumb and optimized MPA
networks.

key results: optimization helps (Nobel prize please). Can reduce percent
of species with conservation loss down to 25% pretty quickly if you
prioritize conservation. Rule of thumb designs like targeting fishing
can get there with perfect information and when considering all animals,
but can be pretty far from optimal with imperfect information or when
only considering a subset of animals in the planning process (that's
what all the different points represent). Some rules of thumb are more
sensitive to these decision than others. E.g. rate can get you optimal,
but goes off the rails pretty quickly with bad information.

There is also a pretty clear breakpoint in results between basically
trying for conservation or trying for economics, and not a lot of wiggle
room in between (the two general bands of lines). xx check to what
extent this depends on the way you scale the two objective function
components xx.

```{r, fig.cap="Percent of species with some conservation loss as a function of MPA size. Points are draws from rule-of-thumb scenarios, lines are results of optimization routine across a gradient of conservation weighting."}

```

Key results: Even with optimization, you can basically either be in a
world with 50% or more of species having some level of unintended
conservation consequence, but no economic loss / some economic gain, or
you can be in a world with lower unintended consequences, but economic
losses. But, even rule of thumb approaches can get you to some places
with relatively little unintended consequences at relatively minimal
economic loss.

```{r, fig.cap="Tradeoff between unitendedn consequences and fisheries revenue. Filled polygons in background represent rule-of-thumb results, with redder colors indicating higher density of points. Points are results of optimization, with color indicating conservation weighting in optimization. Black diagonal line shows 1:1 relationship"}



```

Key results: Presence of and magnitude of unintended consequences pretty
sensitive to design descisions. On a macro scale though far less so. In
terms of change in total SSB/SSB0 (metric of biodiversity, i.e. maxed
out when all species are unfished) and change in revenue, there is
basically a pretty simple linear tradeoff between biodiversity and
revenue, with a slope less than negative 1, meaning that 1% increase in
biodiversity comes at a greater than 1% loss in fishery revenue. This
stacks up pretty nicely with @rassweiler2014., where value of
optimization was low when overfishing of valuable species isn't
prevalent, as is the case here. So, design and optimization can help a
lot with the more "micro" level outcomes (e.g. making sure that
biodiversity gaines aren't coming from massive gains in one species to
detriment of other), but on a macro level pretty insensitive in this
case study.

```{r, fig.cap="Tradeoff frontier between revenue and biodiversity xx across different rule odf thumb design strategies. Points indicate role of thumb results, lines are results of optimized networks across a gradient of conservation weighting."}

```

# Discussion

motivated by but not limited to pelagic ecosystems

**what is the story**

-   Unintended conservation consequences of MPAs resulting from
    technical interactions have been discussed in the literature, but
    almost exclusively in the context of highly tactical modeling or
    localed empirics. We conduct one of the first large-scale
    assessments of the potential of this problem across a range of
    scenarios

-   We find that these unintended consequences can be common outside of
    "canonical" MPA models *a la* hastings etc.

-   But, good news is magnitude of unintended consequences were usually
    small, and total benefits almost always outweigh individual species
    losses (again no probability of states though)

-   Some design strategies are better at avoiding unintended
    consequences than others, but sensitive to data quality.
    Optimization can help, but even with optimization unintended
    consequences can happen.

-   Under the levels of fishing mortality in the WCPO, optimization and
    design can help reduce tradeoffs between fisheries revenue and
    unintended consequences than others, but macro-level tradeoffs
    between biodiversity and fishery revenue are very insensitive to
    design choices, with a pretty clear "increasing biodiversity through
    MPAs reduces fishery revenues".

    -   xx This is clearly hyper-sensitive to the one thing we don't
        iterate over much here: how overfished are the target species.
        At the moment "they're not". But, this runs fast enough, is it
        worth running a hypothetical case study where they are just to
        illustrate how things change?

# Conclusions

-   Unintended conservation outcomes of MPAs totally possible outside of
    "consider a homogenous fish blob" models

-   But, usually small, and net conservation benefits outweigh
    conservation costs

-   Should pay particular attention to potential though when species
    near tipping points are of concern: hard to imagine those not being
    part of the prioritization but something to consider.

-   Other interesting things.

# References

::: {#refs}
:::

# Scraps

Thinking through the "naive" MPA design strategies to compare against
the optimized outcomes. Thoughts at the moment below. These can be run
in a few seconds each, so we cab be be pretty liberal in coming up with
combinations if there are other ideas people want to try, but also if
some of these seem pointless can drop as well! 6:11 Design strategies 1.
Assign protection by SSB / SSB0 weighted abundance in space (protect the
highest concentrations of the most threatened species) 2. Assign
protection by bycatch rates (distribution of biomass bycatch / biomass
target) 3. Assign protection by targeting fishing effort 4. Assign
protection by avoiding fishing effort Knowledge states 1. Degrees of
accuracy over the thing being measured (i.e. error in the design
strategy) 2. Number of species considered (e.g. design aronnd only 2
bycatch species instead of all) 3. über naive: just close off random
patches up to a specified size We then plot these points in objective
function space and compare to the efficiency frontier / consider
frequency and magnitude of conservation and economic tradeoofs The
question all this is answering: how sensitive are conservation and
economic outcomes to design strategy / uncertainty when there are
technical interactions and heterogeneity in habitat? The first stage is
teeing up the idea that beyond conservation / yield tradeoffs, as you
increase habitat heterogeneity with technical interactions you increase
the potential for conservation tradeoffs. But, the degree is going to be
super context dependent, so to dive in a bit we then focus on this case
study of the particular species / habitat overlap in the WCPO. Further
work might reveal more generalities in this, but htis is basically
getting a sense for the potential magnitude of the challenge.

XX stress a key difference: may other papers on this are static, in that
they just some up displacement on different layers, rather than
addressing dynamic respones over tim xx

We show.... - sensitivity of both conservation and fishery outcomes to
pelagic realism and design choices - compared to targeted / dynamic
intervention... - value of information would be great...

Contrast to coastal MPA theoretical work - @hastings2017 - @ovando2021 -
@white2010 - not quite sure what to do about points raised in
@fulton2011a - to what extent has this been done...

The paragraph from @fulton2011a that broke me...

" Spatial management was insufficient as the dominant management lever
because it proves quite difficult to strike the balance between
protecting areas that are of a size sufficient to produce any
conservation benefit and providing sufficient fishing grounds to sustain
an econom- ically viable industry (which is not possible in many
fisheries if closures on the order of 70--80% or more are imposed,
Fulton et al. 2007). More- over, Savina etal. (2008) found that effort
displaced by closures potentially undermines any benefits coming from
the zoning. For instance, strong improvements in diversity (by nearly
20%) under heavy fishing pressure occurred if effort outside a marine
protected area (MPA) was controlled. However, without effort control,
there was a significant (\>20%) drop in overall diversity. Fulton et al.
(2007), Savina et al. (2008), Smith et al. (2010) and unpublished work
associated with Horne et al. (2010) all found that while spatial
management appears to be an effective means of satisfying conservation
objectives con- cerning habitat and restoring age-structures of less
mobile species, benefits from closures for individual species depend on
its trophic position, degree of mobility, geographical distribution and
extent of ontogenetic shifts in habitat use and depth (highly mobile or
prey species do not benefit in the way that more site attached or
predatory species do)."

Lots of studies have shown that lots of things matter, can we provide
guidance on which things are priorities?

Where does this fit in context of @smith2021? One important difference
is that they point out differences in effectiveness, where effectiveness
is still almost always positive, since they just measure leatherbacks.
So their fig 5, there are a few cases where leatherback catches actually
go up, but it's rare. So, extending to multiple species and keeping
track off all of them shows it's even harder

How does this contrast with the the large and growing dynamic ocean
management lit? - By definition, with perfect information dynamic $\geq$
static spatial controls - DOM literature is mostly static, tracks
displacement onto static abundance / risk layers - Doesn't allow for
dynamic impacts, for example fleet adjustments based on updated economic
surface, or short-term costs for long-term benefits (can overstate
costs?) - Dynamic model allows for delta between dynamic policies, e.g.
Fmsy vs DOM - analytical models a la hastings have value but do / don't
stand up to realism

Presentation of the model as a MICE model for MSP - reasonable
parameterizable - flexible - fast

Focus of debate has been on fisheries and yields. As MPAs move to
pelagic systems though, we show that conservation outcomes also depend
on correctly identifying ecological state / design decisions etc.

What are the considerations of protection you need to make good
decisions. It's not always obvious how to design things to get the
outcome you want. If we could make the case that something like closing
the high seas is insufficient conservation management without
conservation management inside EEZs / what is the

Need to do an optimization here, calculate the marginal value of added
cells.

What's this about? Pointing out a problem, which to make impactful would
require making the case that there is a problem with current design
strategies. e.g.. "MPAs are often designed like this (simple dispersive
models with constant habitat, or MARXSAN style things), and doing that
does/does not create problems A tactical evaluation. Picking a handful
of high profile MPAs, parameterizing all the knowns as well as possible,
iterating over unknowns, and see what the range of outcomes are.
Cleaner, more impactful, but also opens it up (depending on the results)
to a details fight about very specific modeling choices How plausible
are losses? Maybe we drill down on this a bit. From a purely bycatch
perspective, the usual questions of fisheries tradeoffs aren't on the
table (but can add to the discussion considerations of how much target
fishing you give up). Given that then, while it's possible that
"targeted" interventions would work better (ignoring costs/enforcement,
etc), there's really no reason not to do MPAs. Unless, you actually end
up concentrating fishing pressure outside on populations. So, we
consider how plausible is it for that to really happen. Issue here is
that seems likely to be pretty hand-wavey. Most likely that the vast
majority of simulations will show positive impacts, but have no way of
assigning probabilities to any one state of nature.

The Role of Blue-Water Marine Protected Areas in Bycatch Reduction

What's the question: How effective are Blue-Water MPAs in reducing
fishing mortality of important pelagic bycatch species?

Another angle. There's going to be a lot of unknowns in designing BWMPA.
How much does not fully understanding a few key life history (e.g.
movement) matter, and how much does the role that the species plays in
the broader fleets portfolio matter

Could simulate habitat from a multivariate normal, where you use the
sigma to control how similar rasters are across species

Marine protected areas are increasingly proposed to protect biodiversity
in pelagic systems. However, maby paradigms of mpa design come from
coastal systems. Here we ask, how likely are blue-water MPAs to provide
conservation benefits to bycatch species, and what are the key
uncertainties underlying these outcomes? We begin with a general
discussion of the ways in which traits of open-ocean systems might
influence the conservation performance of BWMPAs. Given the massive
parameter space, we then constrain ourselves to a styulized version of
the western central pacitic ocean to ask, how feasible is it to provide
broad protection of bycatch species through common MPA design paradigms
in the context of this system. Our results demonstrate the critical
importance of XX and XX in designing MPAs in pelagic systems, and
highlight the potential risks of not accounting for XX. Introductionhere

Bycatch is a significant driver of population declines for many pelagic
species Sharks, seabirds, mammals, etc MPAs increasingly turned to as a
solution for conservation problems, etc. MPA science is largely based on
dynamics of coastal systems. Open-ocean ecosystems have very different
dynamics that require dedicated research Burgess et al. (2018) estimated
the reductions in fishing mortality needed to halt declines for a number
of bycatch species Here we ask, what are the key drivers of bycatch
conservation outcomes from MPAs? There are all of these species; you
reduce fishing mortality. You can either do fisheries management, or you
can do MPAs, let's focus in on the MPA side. This could reduce fishing
mortality, or it could do, and it depends on the dynamics of the species
and the fleets. When can MPAs reduce fishing mortality enough for
bycatch species? And where can you go wrong?

It doesn't make any sense to ask "do MPAs help bycatch": the question is
obviously it depends. To quickly make this case, we'll assume that a
species is in need of recovery. We then do the simple mako example to
show that effects of spatial closure depend on the overlap of habitat
and the assumption of fleet dynamics, and provide some empirical
evidence that in particular effort redistribution is not unlikely.

Determining the effect of an MPA network on bycatch then will also
depend on trying to get close to a tactical model. But, this creates a
problem for blue water systems in particular. Can make the argument that
given the centrality of relatively static habitat features in coastal
MPAs "protect the key habitats" might not be a bad strategy from a
conservation strategy. But in highly mobile pelagic systems there (may
be) a lot more variables in play. Where should researchers direct their
attention?

To reduce the parameter space, we constrain ourselves to a system where
1) we know the species involved 2) we have some sense of the relative
depletion and value of the different species and 3) some basic life
history of the species. We then vary commonly unknown parameters,
particularly preferred habitats, spatial catchability, movement rates
etc, and pair these simulations with different MPA designs, and examine
the conservation outcomes for bycatch species.

What i don't like about this

There isn't an "answer" to this: baring something like "changing the
type of density dependence never had any effect greater than XX%", for
anything that on occasion has an impact the answer of "how much" will
almost always depend on what else is going on.

So, it seems like there are two real angles of attack here

Pointing out a problem, which to make impactful would require making the
case that there is a problem with current design strategies. I.e. "MPAs
are often designed like this (simple dispersive models with constant
habitat, or MARXSAN style things), and doing that does/does not create
problems A tactical evaluation. Picking a handful of high profile MPAs,
parameterizing all the knownns as well as possible, iterating over
unknowns, and see what the range of outcomes are. Cleaner, more
impactful, but also opens it up (depending on the results) to a details
fight about very specific modeling choices

Methods

Collect life history for key bycatch species and target species, cite
Bradley et al. pulling together species distribution and threats etc.
Introduce marlin Parameterize marlin runs with "knowns" from life
history of bycatch and target species Highlight key "unknowns" that we
will iterate over, e.g. fleet models, density dependent movement, timing
of density dependence etc Simulate effects of MPAs given reasonably
realistic situations but iterating over key unknowns Evaluate the range
of outcomes and key drivers of MPA effects I'm thinking we try and avoid
a really specific case study but hard not to accidentally do one without
making everything up....

Results Range of likely changes in fishing-induced mortality for the
species selected Key drivers of uncertainty in those outcomes, focusing
on things that switch states from reduction to increase in mortality How
common are conservation tradeoffs between bycatch species? Fishery
impacts

Discussion What does this say about BWMPAs and bycatch? Not going to get
into design decisions here, depending how common tradeoffs are I think
that will be the main focus Highlight any species that frequently
induced tradeoffs Tee up future research MPAs in context of other
management MPAs as risk reduction strategy Etc.
