---
title: "The Role of MPAs in Bycatch Reduction"
author:
    - Daniel Ovando
    - Darcy Bradley
    - Echelle Burns
    - Lennon Thomas
date: "`r Sys.Date()`"
output: 
  bookdown::pdf_document2: default
  bookdown::html_document2: default
editor_options: 
  markdown: 
    wrap: 72
bibliography: references.bib
params:
  run_name: ["v0.5"]
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, include = FALSE)

source(here::here("scripts", "00_setup.R"))

theme_set(marlin::theme_marlin())

```

```{r load-results}
experiments <- read_rds(file = file.path(results_path, "experiements.rds"))

critter_lookup <- read_rds(file = file.path(results_path, 'critter-lookup.rds'))

rough_habitat <- read_rds(file = file.path(results_path, "rough-habitat.rds"))

a = critter_lookup %>%
  left_join(rough_habitat %>% select(species_commonname, bycatch) %>% unique)


scenes <- experiments %>%
  select(xid, data) %>%
  unnest(cols = data) %>%
  select(xid:seasonal_movement) %>%
  unique()


mpa_results <-  experiments %>%
  select(xid, experiment) %>%
  unnest(cols = experiment) %>%
  select(xid,target_fauna, mpa_size, results) %>%
  unnest(cols = results) %>%
  mutate(target_fauna = map_chr(target_fauna, ~paste(.x, collapse = ","))) %>%
  left_join(a, by = c("critter" = "species_sciname")) %>%
  left_join(scenes, by = "xid") %>%
  mutate(bycatch = ifelse(bycatch == TRUE, "Bycatch Species","Target Species"))

mpa_strategy <- c(
  `carcharhinus longimanus` = "MPAs for OWT Shark",
  `prionace glauca,carcharhinus longimanus,isurus oxyrinchus,carcharhinus falciformis` = "MPAs for All Bycatch"
)



ssb_traj_plot <- mpa_results %>%
  filter(variable == "ssb") %>%
  ggplot(aes(
    mpa_size,
    percent_change_ssb0,
    color = species_commonname,
    group = interaction(xid,species_commonname,bycatch)
  )) +
  geom_hline(aes(yintercept = 0)) +
  geom_line(size = 1.5, alpha = 0.25) +
  facet_grid( bycatch~ target_fauna, labeller = labeller(target_fauna = mpa_strategy), scales = "free_y") +
  scale_x_continuous(name = "MPA Size", labels = scales::percent) +
  scale_y_continuous(name = "% of SSB0 Change", labels = scales::percent) +
  scale_color_discrete(name = "Species") +
  scale_linetype(name = '') +
  theme(axis.text = element_text(size = 8),
        axis.title = element_text(size = 14),
        legend.text = element_text(size = 12))



mpa_strategy <- c(
  `carcharhinus longimanus` = "MPAs for OWT Shark",
  `prionace glauca,carcharhinus longimanus,isurus oxyrinchus,carcharhinus falciformis` = "MPAs for All Bycatch"
)


catch_traj_plot <- mpa_results %>%
  filter(variable == "c") %>%
  ggplot(aes(
    mpa_size,
    percent_change,
    color = species_commonname,
    group = interaction(xid,species_commonname,bycatch)
  )) +
  geom_hline(aes(yintercept = 0)) +
  geom_line(size = 1, alpha = 0.25) +
  facet_wrap(bycatch ~ target_fauna, labeller = labeller(target_fauna = mpa_strategy)) +
  scale_x_continuous(name = "MPA Size", labels = scales::percent) +
  scale_y_continuous(name = "% Change in Catch", labels = scales::percent) +
  scale_color_discrete(name = "Species") +
  scale_linetype(name = '') +
  theme(axis.text = element_text(size = 8),
        axis.title = element_text(size = 14),
        legend.text = element_text(size = 12))



mpa_strategy <- c("MPAs for OWT Shark" = "carcharhinus longimanus" ,
                  "MPAs for All Bycatch" = "prionace glauca,carcharhinus longimanus,isurus oxyrinchus,carcharhinus falciformis")

tradeoff_frontier_plot <- mpa_results %>%
  mutate(target_fauna = as.factor(target_fauna)) %>%
  mutate(target_fauna = fct_recode(target_fauna,!!!mpa_strategy)) %>%
  select(mpa_size,
         percent_change,
         species_commonname,
         target_fauna,
         variable,
         xid) %>%
  filter(variable %in% c("c", "ssb")) %>%
  pivot_wider(names_from = "variable", values_from = percent_change) %>%
  ggplot(aes(ssb, c, shape = target_fauna, linetype = target_fauna)) +
  geom_vline(aes(xintercept = 0)) +
  geom_hline(aes(yintercept = 0)) +
  # geom_line() +
  geom_point(aes(color = mpa_size), size = 2) +
  facet_wrap(~ species_commonname, scales = "free") +
  scale_y_continuous(name = "% Change in Catch", labels = scales::percent) +
  scale_x_continuous(name = "% of SSB0 Change", labels = scales::percent) +
  theme(
    axis.text = element_text(size = 8),
    axis.title = element_text(size = 14),
    legend.text = element_text(size = 12),
    legend.position = "top",
    legend.direction = "horizontal"
  ) +
  scale_colour_viridis_b(
    name = "MPA Size",
    labels = scales::percent,
    guide = guide_colorbar(barwidth = unit(9, "lines"))
  ) +
  scale_shape(name = "") +
  scale_linetype(name = "")





a %>% knitr::kable() %>% kableExtra::kable_paper()


# what is interesting
#
#
# proportion stocks benefiting / losing as a functino of MPA size and spatial homogeneity
#

any_ssb_loss_plot <- mpa_results %>%
  mutate(target_fauna = as.factor(target_fauna)) %>%
  mutate(target_fauna = fct_recode(target_fauna,!!!mpa_strategy)) %>%
  filter(variable == "ssb") %>%
  group_by(mpa_size, sigma_centroid, seasonal_movement, target_fauna) %>%
  summarise(percent_improve = mean(percent_change >= 0),
            percent_lose = 1 - percent_improve) %>%
  ggplot(aes(mpa_size, percent_lose, color = factor(sigma_centroid))) +
  geom_line(size = 1.5) +
  geom_hline(aes(yintercept = 0)) +
  scale_color_viridis_d(name = "Heterogeneity in Adult Habitat") +
  scale_x_continuous(labels = scales::percent, name = 'Area in MPA') +
  scale_y_continuous(labels = scales::percent, name = "Sims where MPA reduces SSB") +
  facet_grid(target_fauna~seasonal_movement, labeller = label_both) +
  theme(legend.position = "top")


signif_ssb_loss_plot <- mpa_results %>%
  mutate(target_fauna = as.factor(target_fauna)) %>%
  mutate(target_fauna = fct_recode(target_fauna,!!!mpa_strategy)) %>%
  filter(variable == "ssb") %>%
  group_by(mpa_size, sigma_centroid, seasonal_movement, target_fauna) %>%
  summarise(percent_improve = mean(percent_change >= .1),
            percent_lose = mean(percent_change <= -.1)) %>%
  ggplot(aes(mpa_size, percent_lose, color = factor(sigma_centroid))) +
  geom_line(size = 1.5) +
  geom_hline(aes(yintercept = 0)) +
  scale_color_viridis_d(name = "Heterogeneity in Adult Habitat") +
  scale_x_continuous(labels = scales::percent, name = 'Area in MPA') +
  scale_y_continuous(labels = scales::percent, name = "Sims where SSB Decreaes by more than 10%") +
  facet_grid(target_fauna~seasonal_movement, labeller = label_both) +
  theme(legend.position = "top")

signif_ssb_gain_plot <- mpa_results %>%
  mutate(target_fauna = as.factor(target_fauna)) %>%
  mutate(target_fauna = fct_recode(target_fauna,!!!mpa_strategy)) %>%
  filter(variable == "ssb") %>%
  group_by(mpa_size, sigma_centroid, seasonal_movement, target_fauna) %>%
  summarise(percent_improve = mean(percent_change >= .1),
            percent_lose = mean(percent_change <= -.1)) %>%
  ggplot(aes(mpa_size, percent_improve, color = factor(sigma_centroid))) +
  geom_line(size = 1.5) +
  geom_hline(aes(yintercept = 0)) +
  scale_color_viridis_d(name = "Heterogeneity in Adult Habitat") +
  scale_x_continuous(labels = scales::percent, name = 'Area in MPA') +
  scale_y_continuous(labels = scales::percent, name = "Sims where MPAs increase by more htan 10%") +
  facet_grid(target_fauna~seasonal_movement, labeller = label_both) +
  theme(legend.position = "top")



mean_ssb_traj_plot <- mpa_results %>%
  mutate(target_fauna = as.factor(target_fauna)) %>%
  mutate(target_fauna = fct_recode(target_fauna,!!!mpa_strategy)) %>%
  filter(variable == "ssb") %>%
  group_by(mpa_size, sigma_centroid, seasonal_movement, target_fauna) %>%
  summarise(net_percent_change = weighted.mean(percent_change,ssb0)) %>%
  ggplot(aes(mpa_size, net_percent_change, color = factor(sigma_centroid))) +
  geom_line(size = 1.5) +
  geom_hline(aes(yintercept = 0)) +
  scale_color_viridis_d(name = "Heterogeneity in Adult Habitat") +
  scale_x_continuous(labels = scales::percent, name = 'Area in MPA') +
  scale_y_continuous(labels = scales::percent, name = "% Change SSB") +
  facet_grid(target_fauna~seasonal_movement, labeller = label_both) +
  theme(legend.position = "top")



total_ssb_traj_plot <- mpa_results %>%
  mutate(target_fauna = as.factor(target_fauna)) %>%
  mutate(target_fauna = fct_recode(target_fauna,!!!mpa_strategy)) %>%
  filter(variable == "ssb") %>%
  group_by(mpa_size, sigma_centroid, seasonal_movement, target_fauna) %>%
  summarise(net_change = sum(abs_change)) %>%
  ggplot(aes(mpa_size, net_change, color = factor(sigma_centroid))) +
  geom_line(size = 1.5) +
  geom_hline(aes(yintercept = 0)) +
  scale_color_viridis_d(name = "Heterogeneity in Adult Habitat") +
  scale_x_continuous(labels = scales::percent, name = 'Area in MPA') +
  scale_y_continuous(name = "Net Change in SSB") +
  facet_grid(target_fauna~seasonal_movement, labeller = label_both) +
  theme(legend.position = "top")


signif_bycatch_ssb_loss_plot <- mpa_results %>%
  filter(bycatch == "Bycatch Species") %>%
  mutate(target_fauna = as.factor(target_fauna)) %>%
  mutate(target_fauna = fct_recode(target_fauna,!!!mpa_strategy)) %>%
  filter(variable == "ssb") %>%
  group_by(mpa_size, sigma_centroid, seasonal_movement, target_fauna) %>%
  summarise(percent_improve = mean(percent_change >= 0),
            percent_lose = 1 - percent_improve) %>%
  ggplot(aes(mpa_size, percent_lose, color = factor(sigma_centroid))) +
  geom_line(size = 1.5) +
  geom_hline(aes(yintercept = 0)) +
  scale_color_viridis_d(name = "Heterogeneity in Adult Habitat") +
  scale_x_continuous(labels = scales::percent, name = 'Area in MPA') +
  scale_y_continuous(labels = scales::percent, name = "Sims where MPA reduces SSB") +
  facet_grid(target_fauna~seasonal_movement, labeller = label_both) +
  theme(legend.position = "top")


catch_loss_plot <- mpa_results %>%
  mutate(target_fauna = as.factor(target_fauna)) %>%
  mutate(target_fauna = fct_recode(target_fauna,!!!mpa_strategy)) %>%
  filter(variable == "c") %>%
  group_by(mpa_size, sigma_centroid, seasonal_movement, target_fauna) %>%
  summarise(percent_improve = mean(percent_change >= 0),
            percent_lose = 1 - percent_improve) %>%
  ggplot(aes(mpa_size, percent_lose, color = factor(sigma_centroid))) +
  geom_line(size = 1.5) +
  geom_hline(aes(yintercept = 0)) +
  scale_color_viridis_d(name = "Heterogeneity in Adult Habitat") +
  scale_x_continuous(labels = scales::percent, name = 'Area in MPA') +
  scale_y_continuous(labels = scales::percent, name = "Sims where MPA reduces Catch") +
  facet_grid(target_fauna~seasonal_movement, labeller = label_both) +
  theme(legend.position = "top")

signif_catch_loss_plot <- mpa_results %>%
  mutate(target_fauna = as.factor(target_fauna)) %>%
  mutate(target_fauna = fct_recode(target_fauna,!!!mpa_strategy)) %>%
  filter(variable == "c") %>%
  group_by(mpa_size, sigma_centroid, seasonal_movement, target_fauna) %>%
  summarise(percent_improve = mean(percent_change >= .1),
            percent_lose = mean(percent_change <= -.1)) %>%
  ggplot(aes(mpa_size, percent_lose, color = factor(sigma_centroid))) +
  geom_line(size = 1.5) +
  geom_hline(aes(yintercept = 0)) +
  scale_color_viridis_d(name = "Heterogeneity in Adult Habitat") +
  scale_x_continuous(labels = scales::percent, name = 'Area in MPA') +
  scale_y_continuous(labels = scales::percent, name = "Sims where MPA reduces Catch by more than 10%") +
  facet_grid(target_fauna~seasonal_movement, labeller = label_both) +
  theme(legend.position = "top")


signif_catch_gain_plot <- mpa_results %>%
  mutate(target_fauna = as.factor(target_fauna)) %>%
  mutate(target_fauna = fct_recode(target_fauna,!!!mpa_strategy)) %>%
  filter(variable == "c") %>%
  group_by(mpa_size, sigma_centroid, seasonal_movement, target_fauna) %>%
  summarise(percent_improve = mean(percent_change > .1),
            percent_lose = mean(percent_change < -.1)) %>%
  ggplot(aes(mpa_size, percent_improve, color = factor(sigma_centroid))) +
  geom_line(size = 1.5) +
  geom_hline(aes(yintercept = 0)) +
  scale_color_viridis_d(name = "Heterogeneity in Adult Habitat") +
  scale_x_continuous(labels = scales::percent, name = 'Area in MPA') +
  scale_y_continuous(labels = scales::percent, name = "Sims where MPA incease Catch by >10%") +
  facet_grid(target_fauna~seasonal_movement, labeller = label_both) +
  theme(legend.position = "top")


```


# Introduction

Marine protected areas (MPAs) are a growing part of the marine
conservation toolbox, with 7.4% of the world's oceans currently
protected and, and bodies such as the IUCN calling for 30% protection by
2030. Many of the largest of these MPAs are being implemented in the
form of "blue-water" MPAs largely covering pelagic ecosystems,
characterized by highly mobile species and fishing fleets using a
diverse mix of static and dynamic ocean features. These systems operate
very differently from the coastal areas where modern MPA design
practices have been developed. Pelagic systems are also characterized by
combinations of relatively abundant species targeted by fishing, such as
most tuna stocks, and highly threatened species caught as incidental
"by-catch" in targeted fishing operations, such as oceanic whitetip
sharks. Blue-water MPAs are increasingly being proposed specifically to
benefit these by-catch populations while minimizing losses or where
possible providing benefits to fisheries.

While the effect of MPAs on fisheries yields is a continuously debated
topic, the conservation impacts of MPAs have received less attention of
late, due in part perhaps to the perceived obvious nature of the
problem: if you prevent fishing of a species within an area, there will
be more of that species in that area, and so the conservation benefits
of MPAs will increase with MPA size. \@hastings2017 clearly illustrates
this idea. Using a simple two-species two-patch analytical model, they
derive conditions under which MPAs can solve the "weak-stock" problem in
which a less productive species is incidentally exploited by a fleet
targeting a more valuable and productive stock.

However, as @hastings2017 point out, their results depend on a very
specific set of assumptions, namely that 1) adults are perfectly
sedentary 2) the target and bycatch species overlap perfectly in terms
of habitat xx....The authors suggest that the impact of these
assumptions should be tested in scenarios under which they are unlikely
to apply. We build on \@hastings2017 ..xx sort of... theirs is about
yield and we don't talk about that much yet... though we can xx to
consider the role of MPAs in protecting bycatch species in pelagic
ecosystems.

The core models underpinning MPA science have largely been based on
stylized versions of near-shore reef-style xx what to call that, like
coral / rocky / kelp forest, etc xx systems, where multiple species
often share common habitat, and adults have reasonable site fidelity
while larvae are broadly dispersed. Pelagic ecosystems have vastly
different dynamics than coastal seas [@game2009; @lubchenco2003],
characterized by highly migratory species specialized to take advantage
of highly dynamic ocean features such as eddies currents while also
utilizing static features such as seamounts. Pelagic ecosystems also
support both coastal domestic fishing industries along with highly
mobile international high-seas fishing operations. Given these dynamics,
it remains unclear how many of the lessons from coastal MPA science are
applicable to pelagic ecosystems [\@ovando2021XX].

Pelagic ecosystems present both opportunities and challenges for MPAs.
The highly productive and mobile nature of targeted fisheries in these
systems may mean that highly targeted MPAs can be designed that provide
effective conservation at minimal fishery costs. Strategic protection of
key aggregation locations may enhance depleted pelagic fisheries [xx
that Indian ocean study I can't remember right now]. xx big question
here: how focused do we want this paper to be? is it just about
conservation tradeoffs or lack thereof with bycatch species, or is about
the net conservation and fishery effects of BWMPAs. I'm inclined the
former to tee-up the later xx. Alternatively, if individual bycatch
species do not share common habitat preferences or fishery
vulnerability, protection in one area targeted to benefit one species
may negatively impact another through effort displacement [e.g.
@abbott2012].

Here, we use a case study motivated by the global expansion of
blue-water MPAs to assess the sensitivity of bycatch populations and
targeted fisheries to MPA design and key life-history variables... and
what else?

-   Case study of WCPO?

-   Comparison to input / output controls

-   Comparison to DOM?

Our goal here is not to provide tactical recommendations on the design
of blue-water MPAs, but rather to provide a strategic evaluation of the
sensitivity of bycatch populations and targeted fisheries to key
unknowns and design choices in MPA planning.

<!-- The outcome of MPAs for fishing and conservation are highly sensitive to -->

<!-- a wide range of economic and ecological factors. In order to constrain -->

<!-- the possible parameter space, we tune a simulation model based around -->

<!-- best available knowledge of the species and fleets operating in the -->

<!-- areas of the western central pacific ocean (WCPO) while allowing key -->

<!-- unknowns to vary, in particular the nature of adult and larval movement, -->

<!-- the degree of habitat overlap between different species, the economic -->

<!-- incentives facing fishing fleets, and the design strategy used in MPA -->

<!-- placement XX. Most conventional MPA design models are either static, -->

<!-- single species, or both. We present a dynamic, multi-species simulation -->

<!-- model that allows us to cohesively assess the impact of MPAs on a suite -->

<!-- of by-catch and targeted species collectively affected by multiple -->

<!-- fishing fleets.  -->

<!-- We show - something about how when species are homogeneous in space, -->

<!-- conservation gains are basically guaranteed except for constant catch -->

<!-- scenearios, which need to be considered in tuna-type scenarios - When we -->

<!-- introduce heterogeneity in stock structure roughly reflective of -->

<!-- distributions in the WCPO we do / do not begin to see tradeoffs in -->

<!-- conservation outcomes among bycatch species - quantify if there is a -->

<!-- concave function to this - Specific design paradigms are / are not more -->

<!-- / less resilient to this issue - Compare tradeoffs between MPA designs -->

<!-- that minimze internal bycatch tradeoffs (there have to be none if MPA = -->

<!-- 100%) and fishery outcomes -->

<!-- We show that MPAs increased biomass of by-catch species in XX% of -->

<!-- simulations, and decreased them in XX% of simulations. On average a 1% -->

<!-- increase in by-catch species resulted in a XX% change in fishery -->

<!-- catches... Outcomes were most sensitive to XX and XX, and least -->

<!-- sensitive to XX and XX. We do not suggest that these same results would -->

<!-- hold for other systems, but rather demonstrate the importance / -->

<!-- resilience of MPAs to key uncertain variables. In particular for pelagic -->

<!-- ecosystems, our work demonstrates the importance of considering the -->

<!-- collective impacts of space, vulnerability, life history, and fleet -->

<!-- dynamics in MPA design for conservation. -->

# Methods

@hastings2017 employed a simply model to derive conditions under which
MPAs would benefit bycatch species. However they also pointed out that
even small increases in model complexity quickly make analytical
solutions to the effects of MPAs intractable. As such, we based our
analysis on a spatially explicit, multi-species, multi-fleet,
bio-economic model we call
[`marlin`](https://github.com/DanOvando/marlin). Models of MPAs span a
wide range of complexity, from simple analytical tools to end-to-end
ecosystem models [@fulton2015]. `marlin` follows a "models of
intermediate complexity" (MICE) philosophy [@plagányi2014] targeted
specifically around the role of technical interactions across species.
Many MPA models model the impacts of MPAs on a single-species level [@ovando2021XX]. This can be useful for for example examining the ways in which species-specific life history traits interact with fleet dynamics to influence MPA outcomes. However, bycatch problems are driven by technical interactions, in which fishing effort for a target species incidentally captures another species xx how do we want to talk about this in here? Like usually incidentally, but clearly sometimes people try and catch sharks too xx. `marlin` accounts for this by allowing the user to simulate an arbitrary number of fleets capable of catching an arbitrary number of species in a spatial area. Different fleets can place different economic value on different species, while also having species-specific selectivity functions. This allows us to simulate bycatch dynamics of MPAs in pelagic ecosystems. 

## `marlin`

Lots of equations. 

## Simulation Experiments

Given the complexity of the operating model used here, a complete sensitivity analysis across all parameters is not practical. Instead, we focus on the role of XX key attributes of pelagic ecosystems: the degree of spatial overlap between populations, the presence of seasonal migrations, and the potential for ontogenetic shifts in habitat use. 

### Habitat Overlap

We simulate habitat overlap through two parameters: the standard deviation of the habitat centroid for each species $\sigma_c$, and the the dispersion of the species around its centroid $\sigma_d$. When $\sigma_c$ is low, all species share a common habitat. As $\sigma_c$ inscreases, the variance in the centroid of each of the simulated species increases. 

We simulate seasonal movement by breaking the year into two seasons. We then set the adult habitat in season two as the negative inverse of the habitat in season one

We simulate ontogenetic movement by allowing for separate habitat layers for recruits and adults. We then simulate a scenario where recruitment habitat is the negative inverse of adult habitat, such that recruits primarily arive in one area of the system, and then move to another area as they grow. 

etc. 

## comparison to...

For each scenario, could then compare that MPA to the amount of yield
lost / bycatch gained by constraining effort / catch...

# Results


## Conservation Outcomes



```{r traj-plot, include=TRUE, fig.cap="Trajectoris of percent change in spawning stock biomass as a function of MPA size. Color indicates separate species, each line is a different simulation."}

ssb_traj_plot + 
  guides(color = FALSE)


```
```{r ssb-plot,fig.cap="Percent of simulations that resulted in any decrease in spawning stock biomass as a function of seasonal movement, habitat heterogeneity, and MPA design.", include=TRUE}
any_ssb_loss_plot
```


```{r signif-ssb-plot, fig.cap= "Percent of simulations that resulted in decreases of 10% or more in spawning stock biomass as a function of seasonal movement, habitat heterogeneity, and MPA design.", include=TRUE}
signif_ssb_loss_plot
```


## Yield Outcomes



```{r catch-plot,fig.cap="Percent of simulations that resulted in any decrease in catch as a function of seasonal movement, habitat heterogeneity, and MPA design.", include=TRUE}
catch_loss_plot
```

```{r signif-catch-plot, fig.cap= "Percent of simulations that resulted in decreases of 10% or more in catch as a function of seasonal movement, habitat heterogeneity, and MPA design.", include=TRUE}
signif_catch_loss_plot
```

## Value of Information

I think this would be one of the easiest and interesting: how good does you habitat knowledge have to be to minimize tradeoffs?


## Tradeoffs

```{r tradeoff-plot, fig.cap= "Tradeoff frontier of catch and spawning stock biomass as a function of MPA size and MPA design. Each point represents a different simulation", include=TRUE}
tradeoff_frontier_plot
```


## Sensitivy to...


Summary (hopefully statistical) of sensitivity of outcome to different parameters? But hard to know if there's really the sample size for that

## Case Study Of...



# Discussion

XX stress a key difference: may other papers on this are static, in that
they just some up displacement on different layers, rather than
addressing dynamic respones over tim xx

We show....
  - sensitivity of both conservation and fishery outcomes to pelagic realism and design choices
  - compared to targeted / dynamic intervention... 
  - value of information would be great... 
  
**what is novel????**
  
Contrast to coastal MPA theoretical work
  - @hastings2017
  - @ovando2021
  - @white2010
  - not quite sure what to do about points raised in @fulton2011a
    - to what extent has this been done...
    
The paragraph from @fulton2011a that broke me...

"
Spatial management was insufficient as the dominant management lever because it proves quite difficult to strike the balance between protecting areas that are of a size sufficient to produce any conservation benefit and providing sufficient fishing grounds to sustain an econom- ically viable industry (which is not possible in many fisheries if closures on the order of 70–80% or more are imposed, Fulton et al. 2007). More- over, Savina etal. (2008) found that effort displaced by closures potentially undermines any benefits coming from the zoning. For instance, strong improvements in diversity (by nearly 20%) under heavy fishing pressure occurred if effort outside a marine protected area (MPA) was controlled. However, without effort control, there was a significant (>20%) drop in overall diversity. Fulton et al. (2007), Savina et al. (2008), Smith et al. (2010) and unpublished work associated with Horne et al. (2010) all found that while spatial management appears to be an effective means of satisfying conservation objectives con- cerning habitat and restoring age-structures of less mobile species, benefits from closures for individual species depend on its trophic position, degree of mobility, geographical distribution and extent of ontogenetic shifts in habitat use and depth (highly mobile or prey species do not benefit in the way that more site attached or predatory species do)."

Lots of studies have shown that lots of things matter, can we provide guidance on which things are priorities?

Where does this fit in context of @smith2021? One important difference is that they point out differences in effectiveness, where effectiveness is still almost always positive, since they just measure leatherbacks. So theri fig 5, there are a few cases where leatherback catches actually go up, but it's rare. So, extending to multiple species and keeping track off all of them shows it's even harder

How does this contrast with the the large and growing dynamic ocean management lit?
  - By definition, with perfect information dynamic $\geq$ static spatial controls
  - DOM literature is mostly static, tracks displacemento onto static abundance / risk layers
  - Doesn't allow for dynamic impacts, for example fleet adjustments based on updated economic surface, or short-term costs for long-term benefits (can overstate costs?)
  - Dynamic model allows for delta between dynamic policies, e.g. Fmsy vs DOM
    - analytical models a la hastings have value but do / don't stand up to realism

Presentation of the model as a MICE model for MSP
  - reasonable parameterizable
  - flexible
  - fast
  

Focus of debate has been on fisheries and yields. As MPAs move to pelagic systems though, we show that conservation outcomes also depend on correctly identifying ecological state / design decisions etc. 
  

# References

::: {#refs}
:::

## Figure Ideas

Are there or are there not tradeoffs in conservation and fisheries
yields

What are the explicit considerations of

Interesting considerations we do biodiversity MPAs we define
biodiversity as system diversity and we put all the species in there,
threatened bycatch species become the the focus and stated objective is
ramp up the value of bycatch species . There might be instances where
just the biodiversity base actually gets you a better outcome than .
There might be overlap and not.

Dig into that mismatch

We can explore that mismatch between is the issue q and is the issue of
spatial habitat

### Fig.1 Motivating figure - basically something like the shortfin mako example

### Fig.2 Case study figure for WCPO: shows degree of overlap in species abundance in space and time, as well as a platform for some examples of MPA networks designed around different paradigms

### Fig.3 Tradeoff plots across species, though will need to put some thought into visualizing across species. But something to convey the idea of whether there are tradeoffs among species in the case study of the WCPO high seas style system. Interestingly, you know that if you closed 100% all benefit, so doing this along a gradient is interesting, are there tradeoffs at low sizes that disappear with higher? Actually you might expect a dome-shaped tradeoff, where there are no tradeoffs at low size since little displacement, largest tradeoffs in the middle where you maximize displacement but still haven't maximized protection, and then no tradeoffs at the extreme where everything is protected. If you have a small enough number of species, say less than 10, then can do color / shape by species, mpa size on x axis, change in biomass on y axis, and connected lines for design paradigms. If not too busy, would be good to have a counterfactual where everything is the same except species are all homogeneously distributed

### Fig 4. Similar idea but, now bringing fishing in there. Maybe collapsing down a few axes, so you do like percent increase in total bycatch by total fishing, or some measure of collective benefit / loss, as a function of MPA size. Maybe highlight inflection points that seem to minimize fishery losses while also minimizing tradeoffs with bycatch

### Fig 5 + Figures showing a few select "what if the world were different", i.e. how important is movement rate, DD form, etc.

# Scraps

What are the considerations of protection you need to make good
decisions. It's not always obvious how to design things to get the
outcome you want. If we could make the case that something like closing
the high seas is insufficient conservation management without
conservation management inside EEZs / what is the

Need to do an optimization here, calculate the marginal value of added
cells.

What's this about? Pointing out a problem, which to make impactful would
require making the case that there is a problem with current design
strategies. e.g.. "MPAs are often designed like this (simple dispersive
models with constant habitat, or MARXSAN style things), and doing that
does/does not create problems A tactical evaluation. Picking a handful
of high profile MPAs, parameterizing all the knowns as well as possible,
iterating over unknowns, and see what the range of outcomes are.
Cleaner, more impactful, but also opens it up (depending on the results)
to a details fight about very specific modeling choices How plausible
are losses? Maybe we drill down on this a bit. From a purely bycatch
perspective, the usual questions of fisheries tradeoffs aren't on the
table (but can add to the discussion considerations of how much target
fishing you give up). Given that then, while it's possible that
"targeted" interventions would work better (ignoring costs/enforcement,
etc), there's really no reason not to do MPAs. Unless, you actually end
up concentrating fishing pressure outside on populations. So, we
consider how plausible is it for that to really happen. Issue here is
that seems likely to be pretty hand-wavey. Most likely that the vast
majority of simulations will show positive impacts, but have no way of
assigning probabilities to any one state of nature.

The Role of Blue-Water Marine Protected Areas in Bycatch Reduction

What's the question: How effective are Blue-Water MPAs in reducing
fishing mortality of important pelagic bycatch species?

Another angle. There's going to be a lot of unknowns in designing BWMPA.
How much does not fully understanding a few key life history (e.g.
movement) matter, and how much does the role that the species plays in
the broader fleets portfolio matter

Could simulate habitat from a multivariate normal, where you use the
sigma to control how similar rasters are across species

Marine protected areas are increasingly proposed to protect biodiversity
in pelagic systems. However, maby paradigms of mpa design come from
coastal systems. Here we ask, how likely are blue-water MPAs to provide
conservation benefits to bycatch species, and what are the key
uncertainties underlying these outcomes? We begin with a general
discussion of the ways in which traits of open-ocean systems might
influence the conservation performance of BWMPAs. Given the massive
parameter space, we then constrain ourselves to a styulized version of
the western central pacitic ocean to ask, how feasible is it to provide
broad protection of bycatch species through common MPA design paradigms
in the context of this system. Our results demonstrate the critical
importance of XX and XX in designing MPAs in pelagic systems, and
highlight the potential risks of not accounting for XX. Introductionhere

Bycatch is a significant driver of population declines for many pelagic
species Sharks, seabirds, mammals, etc MPAs increasingly turned to as a
solution for conservation problems, etc. MPA science is largely based on
dynamics of coastal systems. Open-ocean ecosystems have very different
dynamics that require dedicated research Burgess et al. (2018) estimated
the reductions in fishing mortality needed to halt declines for a number
of bycatch species Here we ask, what are the key drivers of bycatch
conservation outcomes from MPAs? There are all of these species; you
reduce fishing mortality. You can either do fisheries management, or you
can do MPAs, let's focus in on the MPA side. This could reduce fishing
mortality, or it could do, and it depends on the dynamics of the species
and the fleets. When can MPAs reduce fishing mortality enough for
bycatch species? And where can you go wrong?

It doesn't make any sense to ask "do MPAs help bycatch": the question is
obviously it depends. To quickly make this case, we'll assume that a
species is in need of recovery. We then do the simple mako example to
show that effects of spatial closure depend on the overlap of habitat
and the assumption of fleet dynamics, and provide some empirical
evidence that in particular effort redistribution is not unlikely.

Determining the effect of an MPA network on bycatch then will also
depend on trying to get close to a tactical model. But, this creates a
problem for blue water systems in particular. Can make the argument that
given the centrality of relatively static habitat features in coastal
MPAs "protect the key habitats" might not be a bad strategy from a
conservation strategy. But in highly mobile pelagic systems there (may
be) a lot more variables in play. Where should researchers direct their
attention?

To reduce the parameter space, we constrain ourselves to a system where
1) we know the species involved 2) we have some sense of the relative
depletion and value of the different species and 3) some basic life
history of the species. We then vary commonly unknown parameters,
particularly preferred habitats, spatial catchability, movement rates
etc, and pair these simulations with different MPA designs, and examine
the conservation outcomes for bycatch species.

What i don't like about this

There isn't an "answer" to this: baring something like "changing the
type of density dependence never had any effect greater than XX%", for
anything that on occasion has an impact the answer of "how much" will
almost always depend on what else is going on.

So, it seems like there are two real angles of attack here

Pointing out a problem, which to make impactful would require making the
case that there is a problem with current design strategies. I.e. "MPAs
are often designed like this (simple dispersive models with constant
habitat, or MARXSAN style things), and doing that does/does not create
problems A tactical evaluation. Picking a handful of high profile MPAs,
parameterizing all the knownns as well as possible, iterating over
unknowns, and see what the range of outcomes are. Cleaner, more
impactful, but also opens it up (depending on the results) to a details
fight about very specific modeling choices

Methods

Collect life history for key bycatch species and target species, cite
Bradley et al. pulling together species distribution and threats etc.
Introduce marlin Parameterize marlin runs with "knowns" from life
history of bycatch and target species Highlight key "unknowns" that we
will iterate over, e.g. fleet models, density dependent movement, timing
of density dependence etc Simulate effects of MPAs given reasonably
realistic situations but iterating over key unknowns Evaluate the range
of outcomes and key drivers of MPA effects I'm thinking we try and avoid
a really specific case study but hard not to accidentally do one without
making everything up....

Results Range of likely changes in fishing-induced mortality for the
species selected Key drivers of uncertainty in those outcomes, focusing
on things that switch states from reduction to increase in mortality How
common are conservation tradeoffs between bycatch species? Fishery
impacts

Discussion What does this say about BWMPAs and bycatch? Not going to get
into design decisions here, depending how common tradeoffs are I think
that will be the main focus Highlight any species that frequently
induced tradeoffs Tee up future research MPAs in context of other
management MPAs as risk reduction strategy Etc.
