---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# Something about MPAs and Bycatch

Repo for manuscript exploring the ways in which habitat overlap, fleet spillover, and MPA design strategies can interact to affect conservation and fishery outcomes of MPAs, focused on WCPO-ish system. 

starginw thi hastings botsford weakstock species etc. can you hedge around this or not?Build intuition around BWMPAs for weakstocks 

Covariance matrix across species: perfectly correlated to totally uncorrelated and everything in between 

Clean up by setting Hastings and Botsford as a perfect template for this. PNAS? Isolate the life history 


It is clear that if species are mutually exclusive, there is no potential for overlap protection... but where does it fall when overlap is something in between


Hmmm you could use virtual species.. or I guess that's even too complicated

What if you just picked an "environmental" variable, and generated a random matrix

For each species, you then assign a beta to that variable, where when they are all the same the covariance across the species is high. When they are all near zero there is no correlation, when they are opposite signs they are negatively correlated?

Suppose you had two variables actually. One specifies kind of the "default" distributino for all species. The other is an environmental variable, and species can have coefficients that on average range from all 1, evenly split 1 and -1, and near zero


Ah ok what about this. YOu define a set of benchmark species, where those benchmark species are defined by their relationship to a habitat. FOr the remainder of the species, their habitat is on average with noise on a gradient from -1 to 1, where one means they have the same as the benchmark, 0 means they are uncorrlated with the benchmark species, and -1 they are negatively correlated. 

```{r}

library(marlin)
library(tidyverse)

theme_set(marlin::theme_marlin())

resolution <- 20 # resolution is in squared patches, so 20 implies a 20X20 system, i.e. 400 patches 

seasons <- 1

years <- 50

tune_type <- "depletion"

steps <- years * seasons

# for now make up some habitat

base_hab <- expand_grid(x = 1:resolution, y = 1:resolution) %>%
  mutate(habitat =  .05 * x * y) %>% 
  mutate(habitat = as.numeric(scale(habitat)))

 


hab_scenes <- expand_grid(scene = c("friends", "enemies","frenemies"), species = fruit[1:10]) %>% 
  group_by(scene) %>% 
  nest()


generate_hab <- function(species, homogifier = 1,scene,base_hab,sigma_hab = 1, beta_hab = 1){
  
  
  sigma_hab = .2
  
  sigma_block <- 0.1
  
  homogifier <- 0
  
  species <- hab_scenes$data[[1]]
  
  block_a <- species$species[1:5]
  
  block_b <- species$species[6:10]
  
  block_a_betas <- rnorm(length(block_a),1,sigma_block)
  
  block_b_betas <- rnorm(length(block_b), homogifier, sigma_block)
  
  species$betas <- c(block_a_betas, block_b_betas)
  
  species$habitat <- vector(mode = "list", length = nrow(species))
  
  for (i in 1:nrow(species)){
    
    
    base_hab$realized_hab <- rnorm(nrow(base_hab), base_hab$habitat * species$betas[i], sigma_hab)
    
    species$habitat[[i]] <- base_hab
      
    
  }
  

  (species$habitat[[1]] %>% 
    ggplot(aes(x,y,fill = realized_hab))+ 
    geom_tile() + 
    scale_fill_viridis_c()) + 
      (species$habitat[[10]] %>% 
    ggplot(aes(x,y,fill = realized_hab))+ 
    geom_tile() + 
    scale_fill_viridis_c())
  
  
  
}



```

